---
title: "P8105 Homework 3"
author: "Madison Goldrich mpg2166"
output: github_document
---

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = .8,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1

#### Load the data:

```{r}
data("instacart")

instacart =
  instacart |> 
  as_tibble()
```

#### Answer questions about the data: 

The `instacart` dataset is `r nrow(instacart)` rows by `r ncol(instacart)` columns. According to the description on the course website, "The dataset contains 1,384,617 observations of 131,209 unique users, where each row in the dataset is a product from an order." Some key variables include `product_id` which identifies the product, `order_id` which identifies the order in which the product was placed, `reordered` which identifies whether the product has been ordered by the user in the past, among others. In total there are `r instacart |> select(product_id) |>  distinct() |> count()` products found in `r instacart |> select(user_id, order_id) |>  distinct() |>  count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

Below is a table summarizing the number of items ordered from each aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits having by far the most items ordered.

```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle, with aisles ordered by ascending number of items.

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

The next table shows the three most popular items in `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, including the number of times each item is ordered.

```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |> 
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |> 
  knitr::kable()
```

Table showing the mean hour of the day at which Pink Lady Apples and Coffe Ice Cream are ordered on each day of the week. Pink lady apples are generally purchased slightly earlier in the day than coffee ice cream, except on day 5.

```{r}
instacart |> 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |> 
  group_by(product_name, order_dow) |> 
  summarize(mean_hour = mean(order_hour_of_day)) |> 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) |> 
  knitr::kable(digits = 2)
```


## Problem 2

#### Load and clean the data:

```{r}
data("brfss_smart2010")

brfss_smart2010 = 
  brfss_smart2010 |> 
  as_tibble() |> 
  janitor::clean_names() |> 
  rename(
    state = locationabbr, 
    location = locationdesc,
    resp_id = respid) |> 
  filter(
    topic == "Overall Health",
    response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) |> 
  mutate(response = factor(
    response,
    levels = c("Poor", "Fair", "Good", "Very good", "Excellent")))
```

#### Answer questions about the data:

Below are 2 tables that show the states that were observed at 7 or more locations. The first table for the year 2002, with 6 states represented, and the second table for 2010 with 14 states.

```{r}
brfss_smart2010 |> 
  filter(year == 2002) |> 
  distinct(location, .keep_all = TRUE) |> 
  count(state) |> 
  filter(n >= 7) |> 
  knitr::kable()

brfss_smart2010 |> 
  filter(year == 2010) |> 
  distinct(location, .keep_all = TRUE) |> 
  count(state) |> 
  filter(n >= 7) |> 
  knitr::kable()
```

The below code onstructs a dataset `bfss_smart2010_excellent` that is limited to Excellent responses, and contains year, state, and a variable `mean_data_value` that averages the data_value across locations within a state. 

```{r}
brfss_smart2010_excellent =
  brfss_smart2010 |> 
  filter(response == "Excellent") |> 
  group_by(year, state) |> 
  summarize(
    mean_data_value = mean(data_value)
  )
```

This "spaghetti" plot shows the average value over time within a state.

```{r}
brfss_smart2010_excellent |> 
  ggplot(aes(x = year, y = mean_data_value, color = state)) +
  geom_line(aes(group = state)) +
  labs(title = "Spaghetti plot of mean data value over time by state")
```

The below plot shows the disribution of `data_value` by response level among locations in New York state in the years 2006 and 2010.

```{r}
brfss_smart2010 |> 
  filter(
    state == "NY",
    year %in% c(2006, 2010)) |> 
  ggplot(aes(x = data_value, fill = response)) +
  geom_density(alpha = .4) +
  facet_grid(. ~ year) +
  labs(title = "Distribution of data_value by response in NY state")
```


## Problem 3

#### Load, tidy, merge, and organize the datasets:

```{r}
nhanes_dem =
  read_csv("data/nhanes_covar.csv", skip = 4) |> 
  janitor::clean_names() |> 
  mutate(
    sex = case_match(
      sex,
      1 ~ "Male",
      2 ~ "Female"),
    education = case_match(
      education,
      1 ~ "Less than High School",
      2 ~ "High School Equivalent",
      3 ~ "More than High School"),
    education = factor(
      education,
      levels = c("Less than High School", "High School Equivalent", "More than High School")
    ))

nhanes_accel =
  read_csv("data/nhanes_accel.csv") |> 
  janitor::clean_names()

nhanes =
  full_join(nhanes_dem, nhanes_accel) |> 
  drop_na() |> 
  filter(age >= 21)
```

Below is a table that shows the number of men and women in each education category. The values are fairly evenly distributed in the Less than High School and More than High School categories, but there are a noticeable amount more men than women in the High School Equivalent category.

```{r}
nhanes |> 
  group_by(education, sex) |> 
  count(sex) |> 
  pivot_wider(
    names_from = sex,
    values_from = n
  ) |> 
  knitr::kable()
```

The boxplot below shows the distribution of age by sex within each education category. All 6 boxplots have very similar age ranges. The median age for both sexes is visibly lower in the More than High School category compared with the other two categories. Distributions between sexes appear more aligned in the Less than High School and More than High School categories, whereas ages are clustered at higher values for females than males in the High School Equivalent category.

```{r}
nhanes |> 
  ggplot(aes(x = sex, y = age)) +
  geom_boxplot() +
  facet_grid(. ~ education) +
  labs(title = "Age distributions by sex and education category")
```

The below plot shows total activity vs. age comparing by sex within each education level category. The `geom_smooth` lines show that total activity tends to decrease with age, though it does fluctuate greatly. Total activity levels appear to decrease by education level. 

```{r}
nhanes |> 
  pivot_longer(
    min1:min1440,
    names_to = "minute",
    values_to = "accel"
  ) |> 
  group_by(seqn) |> 
  mutate(
    total_activity = sum(accel)
  ) |> 
  ggplot(aes(x = age, y = total_activity, color = sex)) +
  geom_point(alpha = .7) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ education) +
  labs(title = "Age vs. Total Activity (by Sex and Education Level)")
```

The below three-panel plot shows the 24-hour activity time courses for each education level by sex. Knowing that minute = 1 occurs at 12:01 AM, it makes sense that the general pattern shows a drop-off in activity over the following ~4 hours. Activity sharply picks up from around minute 250 (around 4 AM) until about minute 600 (10 AM) for all groups, remains fairly steady until about minute 1100 (6:30 PM), then declines again until midnight. This makes logical sense! 

```{r}
nhanes |> 
  pivot_longer(
    min1:min1440,
    names_to = "minute",
    names_prefix = "min",
    values_to = "accel"
  ) |> 
  mutate(minute = as.integer(minute)) |> 
  group_by(education, sex, minute) |> 
  summarize(
    mean_accel = mean(accel)
  ) |> 
  ggplot(aes(x = minute, y = mean_accel, color = sex)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ education) 
```



